Prestige as a Driving Force in Cultural Transmission

### 2022-05-29 ms/PrestigeBias.tex
revised entire fixation prob section

\subsection*{Continuous trait}
We follow the model of \citet{evolutionBook}, assuming only oblique transmission of the trait and omitting the indirect trait in order to reduce model complexity. 
We consider a population of $N$ individuals, described by a single trait with a continuous value.
Each generation, $N$ naive individuals, or copiers, choose an individual from the previous generation, or role-models, from which they will copy their trait. Similar to a Wright–Fisher model, we assume non-overlapping generations such that the entire population is replaced in each generation.
The population at time $t$ can be described by $\vec{A}(t)=\big(A_{1}(t), \ldots, A_{N}(t)\big)$ where $A_{i}(t)$ is trait value of individual $i$ at time $t$. We assume the initial population is drawn from a standard normal distribution, $\vec{A}(0) \sim N(0,1)$ .
Cultural transmission is modeled by a function $F$ such that 
\begin{equation}\label{eq:transmission}
A_{i}(t+1) = F_i(\vec{A}(t)) \;.
\end{equation}

\paragraph{Success bias.}
\citet[Ch.8, p.247-249]{evolutionBook} describe a blended transmission algorithm by defining $F$ as a weighted average of the traits of all role-models, 
\begin{equation}\label{eq:boydF}
F_i(\vec{A}) = \sum_{j=1}^N G_{i,j}\cdot A_{i,j} \;, 
\end{equation}
where $G_{i,j}$ is the success bias of role-model $j$ in the eyes of copier $i$,
\begin{equation}\label{eq:boydG}
G_{i,j} = \frac{\beta(A_{i,j})}{\sum_{k=1}^{N} \beta(A_{i,k})} \;,
\end{equation}
$A_{i,j}$ is the absolute trait value copier $i$ estimates for role-model $j$ with some error $e_i \sim N(0,\eta^2)$,
\begin{equation}\label{eq:relativeTrait}
A_{i,j} = A_j + e_i,
\end{equation}
and $\beta(\cdot)$ is the bias function that quantifies the success bias of a role-model,
\begin{equation}\label{eq:success_bias}
\beta(A_{i,j}) = b \cdot \exp{\Big(-\frac{(A_{i,j} - \hat{A})^2}{2J}\Big)},
\end{equation} 
with $\hat{A}$ as the arbitrary optimal trait value, and $J$ and $b$ as parameters that control the bias strength.
Therefore, $G_{i,j}$ is a relative success score that copier $i$ assigns to role-model $j$, equivalent to relative fitness in evolutionary-genetic transmission models.

@@

After finding that the DM distribution is a good approximation of the (within-generations) role-model choice process, we turn our attention to the (between-generations) evolutionary dynamics.
We focus on the fixation probability and fixation time of an advantageous phenotype, similar to analyses in population-genetic models \citep{kimura,kimura_average,otto_fixation}.

We are mainly interested in the effect of the bias weight ($\alpha$), which determines the relative effects of success and influence on prestige bias.
For simplicity, we do not include estimation error in this analysis.
As shown above, transmission in our model is approximately DM distributed with the parameters $\sim DM(\alpha'X,(N-X)\alpha'\beta(A))$ based on \cref{cor:dirichlet} and \cref{eq:copier_choose}.

\begin{remark}[Effective selection coefficient]
In population genetics a selection coefficient is a measure of differences in relative fitness. Selection coefficients are central to the quantitative description of evolution, since fitness differences determine the change in genotype frequencies attributable to selection.
Effective selection coefficient is the equivalent in a model to the selection coefficient in a Wright-Fisher model.
\end{remark}

\begin{result}[Effective selection coefficient]\label{res:selection_coef}
$1-\beta(A)$ is equivalent to the selection coefficient $s$ in the diffusion-equation approximation of the a classic Wright-Fisher model that approximate the fixation probability and fixation time of an advantageous allele.
\end{result}

\begin{proof}
Let $x$ be the frequency of type $\hat{A}$ in the population with $N$ individuals. Let $X$ be the number of individuals of type $\hat{A}$ so $x=X/N$. $X'$ is the number of individuals with $\hat{A}$ in the next generation and $x'$ their frequency.
By definition $\beta(\hat{A})=1$, and for simplicity we'll denote $\beta(A)=\beta$ ($\beta<1$).

The expected number of individuals of a DM distribution is:
\begin{equation}
E[X'] = N  \frac{\alpha_1}{\alpha_1+\alpha_2},
\end{equation}
where $\alpha_1 = \alpha' X$ and $\alpha_2 = \alpha'(N-X)\beta$, from  \cref{eq:binary-model}.
We want to use frequencies instead of quantities to follow Durret's process so:
\begin{equation}
E[x'] = E[\frac{X'}{N}] = \frac{1}{N}E[X']
\end{equation}
Putting it together we get:
\begin{equation}
\begin{split}
E[x'] &= \frac{1}{N}N\frac{\alpha' xN}{\alpha' xN + \alpha' N (1-x)\beta}\\
	 &= \frac{x}{x + (1-x)\beta}
\end{split}
\end{equation}

which is identical to the equation in the top of page 253, chap 7.2 in \citet{durret}. We therefore use the same approximation and say that:
\begin{equation}
\begin{split}
E[x'] =& \frac{x}{x + (1-x)\beta} = \frac{x}{x + (1-x)(1-s)} =\\
 &= x + x(1-x)s + o(s)\\
  &= x + x(1-x)(1-\beta) + o(1-\beta)
\end{split}
\end{equation}

By definition, $x$ is constant, so $E[x] = x$. We continue to calculate $E[x'-x]$:
\begin{equation}\label{eq:expec_freq}
E[x'-x] = E[x'] - E[x] = x(1-x)(1-\beta) + o(1-\beta)
\end{equation}
where when substituting $1-\beta$ with $s$, we get the same result as \citet{durret} which is the desired result.
\end{proof}

\begin{remark}
The effective size of a population, $N_e$, determines the rate of change in the composition of a population caused by genetic drift, which is the random sampling of genetic variants in a finite population. Ne is crucial in determining the level of variability in a population, and the effectiveness of selection relative to drift.
\end{remark}

\begin{result}[Effective population size]\label{res:effective_population}
$N_e=\alpha N + (1-\alpha)$, where $Ne$ is the effective population size of the binary model.
\end{result}

\begin{proof}
The variance of a DM distribution is:
\begin{equation}
V(X') = N\frac{\alpha_1}{\alpha_1+\alpha_2}(1-\frac{\alpha_1}{\alpha_1+\alpha_2})
(\frac{N + \alpha_1+\alpha_2}{1+\alpha_1+\alpha_2})
\end{equation}
And again, we want to use frequencies so:
\begin{equation}
V(\frac{X'}{N}) = \frac{1}{N^2}V(x')
\end{equation}

Putting it together with our model's notations:
\begin{equation}
V(x') = \frac{1}{N^2}N\frac{x}{x+(1-x)\beta}(1-\frac{x}{x+(1-x)\beta})
(\frac{N + \alpha' xN + \alpha' N(1-x)\beta}{1 + \alpha' xN + \alpha' N(1-x)\beta}) 
\end{equation}

Like Durrett, we'll use the zero order of the approximation when $\beta\approx1$,so:
\begin{equation}
\frac{x}{x + (1-x)\beta} \approx x
\end{equation}
and we also use $\beta\approx1$ for the entire variance expression and get:
\begin{equation}
\begin{split}
V(x') & \approx  \frac{1}{N} x(1-x)
(\frac{N + \alpha' xN + \alpha' N - \alpha' xN}{1 + \alpha' xN + \alpha' N - \alpha' xN})\\
&=  x(1-x)(\frac{1 + \alpha'}{1 + \alpha' N}) 
\end{split}
\end{equation}

Again following Durrett we want to calculate:
\begin{equation}\label{eq:var_diff_durret}
V(x'-x) = V(x') - V(x) \approx  x(1-x)(\frac{1 + \alpha'}{1 + \alpha' N})
\end{equation}
because $x$ is a constant so $V(x) = 0$

In our model, $\alpha'$ is the odds ratio of a parameter we called "bias weight", denoted in our model as $\alpha$, so:
\begin{equation}\label{eq:success_ratio}
\alpha' = \frac{\alpha}{1-\alpha}
\end{equation}

Combining \cref{eq:var_diff_durret} and \cref{eq:success_ratio} we get:
\begin{equation}\label{eq:const_var}
\begin{split}
V(x'-x) & \approx x(1-x)(\frac{1 + \frac{\alpha}{1-\alpha}}{1 + \frac{\alpha}{1-\alpha} N})\\
 &= x(1-x)(\frac{\frac{1-\alpha+\alpha}{1-\alpha}}{\frac{1-\alpha+\alpha N}{1-\alpha}})\\
 &= x(1-x)(\frac{1}{1- \alpha(1-N)})\\
  &= x(1-x)(\frac{1}{\alpha N + (1-\alpha)})\\
  &= x(1-x)\frac{1}{N_e}
\end{split}
\end{equation}
\end{proof}

\begin{result}[Fixation probability approximation]
The fixation probability is estimated by:
\begin{equation}\label{eq:kimura_p}
P_{fix} = \frac{1-e^{-2(1-\beta)N_e x}}{1-e^{-2(1-\beta)N_e}}
\end{equation}
where $x$ is the initial frequency of the advantageous phenotype $\hat{A}$.
\end{result}
\begin{proof}
Using our substitute for a selection coefficient, $1-\beta$, and the effective population size $N_e$, we can estimate the fixation probability using the diffusion equation, $E[X'-X]$ (\cref{res:selection_coef}), and $V(X'-X)$ (\cref{res:effective_population}).
\end{proof}

\begin{result}[Fixation time approximation]
\begin{equation}\label{eq:kimura_t}
T_{fix}=\frac{1-P_{fix}}{1-\beta}\int_0^x\frac{e^{2(1-\beta) \xi}-1}{\xi(1-\xi)}d\xi+ \frac{P_{fix}}{1-\beta}\int_x^1\frac{1-e^{-2(1-\beta)(1-\xi)}}{\xi(1-\xi)}d\xi
\end{equation}
\end{result}

\begin{proof}
Again using the diffusion equation and \cref{res:selection_coef,res:effective_population}, we get our result.
the integrals cannot be solved in closed form, so we can only estimate them numerically.
\end{proof}

\paragraph{Numeric validation.}
To validate our math we ran multiple simulations comparing our binary model with the classic Wright-Fisher model, using different $\alpha$ and $\beta$ each time, and using the corresponding values of $s$ and $N_e$ for the WF simulations.
In \cref{fig:var_alpha} we changed $\alpha$ (and $N_e$ accordingly) and used a constant $\beta$.
In \cref{fig:var_beta} we changed $\beta$ and used a constant $\alpha$.
In both cases we can see that the two models behave similarly, and both are approximated well by the Kimura's equations: \cref{eq:kimura_p} and \cref{eq:kimura_t}.

\begin{figure}[h]
    \includegraphics[width=\linewidth]{../figures/binary/kimura_var_alpha.pdf}
  \caption{\textbf{Goodness of fit of Kimura's approximation for varying bias weights.}
  The approximation (black) fits DM simulations results (green) and Wright Fisher simulation results (orange) when using \cref{eq:kimura_p}.
  Fixation probability (a) is limited at approximately $2s$ (blue), like the classic WF model.
  Markers for average values across $5,000$ simulations, error bars for 95\% confidence intervals.
   Here, Population size, $N=1,000$; phenotype values, $\hat{A}=1$, $A=0.7$; success coefficient, $1-\beta=s=0.044$.}
  \label{fig:var_alpha}
\end{figure}

\begin{figure}[h]
    \includegraphics[width=\linewidth]{../figures/binary/kimura_var_beta.pdf}
  \caption{\textbf{Goodness of fit of Kimura's approximation for varying success coefficients.}
   The approximation (black) fits DM simulations results (green) and Wright Fisher simulation results (orange) when using \cref{eq:kimura_p}.
  Markers for average values across $5,000$ simulations, error bars for 75\% confidence intervals.
 Here, Population size, $N=1,000$; phenotype values, $\hat{A}=1$, $A=a\cdot \hat{A}, a \in [0.01,0.99]$; bias weight, $\alpha=0.01$.}
  \label{fig:var_beta}
\end{figure}

After finding good estimations for our model in a constant environment, when the favorable trait is always $\hat{A}$, we want to find an estimation for our model in a changing environment.

For that we will find an expression for the expected and variance of the change in frequency between $t$ generations.
Let $s_t=N(1-\beta_t)$, and $S_n=\sum\limits_{i=1}^n s_i$, where $\beta_t$ is $\beta(A)$ at time/generation $t$.

\begin{result} [Expected difference in frequency in a changing environment]\label{res:ch_expected}
$E[\frac{X_t}{N}-x]\simeq \frac{1}{N}S_tx(1-x)$, where $x$ is the initial frequency of the favorable/invading trait and $X_t$ is the number of individuals with the trait at time $t$.
\end{result}
\begin{result}[Variance of difference in frequency in a changing environment]\label{res:ch_var}
$V(\frac{X_t}{N})\simeq\frac{1}{N_e}tx(1-x)$, where $x$ is the initial frequency of the favorable/invading trait and $X_t$ is the number of individuals with the trait at time $t$.
\end{result}

\begin{proof}
Following the proof of \citet{changeEnv}, we prove by induction. We will prove \cref{res:ch_expected,res:ch_var} together, because the induction assumptions are needed for both.

From \cref{eq:expec_freq} we know that
\begin{equation}\label{eq:ch_1}
\begin{split}
E\left[\frac{X_{t+1}}{N}-\frac{X_t}{N}\bigg|X_t\right] &= \frac{X_t}{N}\left(1-\frac{X_t}{N}\right)(1-\beta_{t+1}) \\
&= \frac{1}{N}\frac{X_t}{N}\left(1-\frac{X_t}{N}\right)s_{t+1}
\end{split}
\end{equation}
Also note that using the definition of $V(y)=E[y^2]-(E[y])^2$
\begin{equation}
\begin{split}
E\left[\frac{X_t}{N}\left(1-\frac{X_t}{N}\right)\right] &= E\left[\frac{X_t}{N}-\left(\frac{X_t}{N}\right)^2\right] \\
&= E\left[\frac{X_t}{N}\right]-E\left[\left(\frac{X_t}{N}\right)^2\right] \\
&= E\left[\frac{X_t}{N}\right] - V\left(\frac{X_t}{N}\right) - \left(E\left[\frac{X_t}{N}\right]\right)^2
\end{split}
\end{equation}

we can now use the induction assumption of $V(\frac{X_t}{N})$ and get
\begin{equation}
\begin{split}
E\left[\frac{X_t}{N}\left(1-\frac{X_t}{N}\right)\right] &\simeq
E\left[\frac{X_t}{N}\right]\left(1-E\left[\frac{X_t}{N}\right]\right)-\frac{1}{N_e}tx(1-x)
\end{split}
\end{equation}
From \cref{eq:ch_1} we know that
\begin{equation}
\begin{split}
E\left[\frac{X_{t+1}}{N}-\frac{X_t}{N}\right] &= \frac{1}{N}s_{t+1}E\left[\frac{X_t}{N}\left(1-\frac{X_t}{N}\right)\right] \\
&\simeq \frac{1}{N}s_{t+1}\left(E\left[\frac{X_t}{N}\right]\left(1-E\left[\frac{X_t}{N}\right]\right) - \frac{1}{N_e}tx(1-x)\right) \\
&\simeq \frac{1}{N}s_{t+1}\cdot E\left[\frac{X_t}{N}\right]\left(1-E\left[\frac{X_t}{N}\right]\right) - \frac{1}{N_e N}s_{t+1}tx(1-x)
\end{split}
\end{equation}
Now we'll omit $O(\frac{1}{Ne\cdot N})$ and get
\begin{equation}\label{eq:ch_2}
E\left[\frac{X_{t+1}}{N}-\frac{X_t}{N}\right] \simeq \frac{1}{N}s_{t+1}\cdot E\left[\frac{X_t}{N}\right]\left(1-E\left[\frac{X_t}{N}\right]\right)
\end{equation}

We'll now look at the induction assumption to see that
\begin{equation}
E\left[\frac{X_t}{N}-x\right]=E\left[\frac{X_t}{N}\right]-E[x]=E\left[\frac{X_t}{N}\right]-x ,
\end{equation}
so using the assumption we get
\begin{equation}
\begin{split}
E\left[\frac{X_t}{N}\right] &\simeq \frac{1}{N} S_t x(1-x)+x \\
1 - E\left[\frac{X_t}{N}\right] &\simeq 1- \frac{1}{N} S_t x(1-x)+x
\end{split}
\end{equation}
we'll use both expressions in \cref{eq:ch_2} and get
\begin{equation}\label{eq:ch_3}
\begin{split}
E\left[\frac{X_{t+1}}{N}-\frac{X_t}{N}\right] &\simeq \frac{1}{N}s_{t+1} \left(\frac{1}{N} S_t x(1-x)+x \right)\left(1- \frac{1}{N} S_t x(1-x)+x \right) \\
&\simeq  \frac{1}{N}s_{t+1}\cdot x(1-x)
\end{split}
\end{equation}
after again omitting $O(\frac{1}{N^2})$ parts of the equation.
To conclude our proof, we see that
\begin{equation}
E\left[\frac{X_{t+1}}{N}-x\right] = E\left[\frac{X_{t+1}}{N}-\frac{X_t}{N}\right] + E\left[\frac{X_t}{N}-x\right]
\end{equation}
so again using the induction assumption, together with \cref{eq:ch_3} we get
\begin{equation}
\begin{split}
E\left[\frac{X_{t+1}}{N}-x\right] &\simeq \frac{1}{N}s_{t+1}\cdot x(1-x) + \frac{1}{N}S_t\cdot x(1-x) \\
& \simeq \frac{1}{N}x(1-x)(S_t + s_{t+1}) \\
& \simeq \frac{1}{N} S_{t+1} x(1-x)
\end{split}
\end{equation}
which proves \cref{res:ch_expected}.

For \cref{res:ch_var}, we'll use a property of variance:
\begin{equation}\label{eq:ch_var}
V\left(\frac{X_{t+1}}{N}\right) = E\left[V\left(\frac{X_{t+1}}{N} \bigg|X_t \right)\right] + V\left(E\left[\frac{X_{t+1}}{N} \bigg|X_t \right]\right)
\end{equation}

>>>>

\subsection*{Continuous trait}
We follow the model of \citet{evolutionBook}, assuming only oblique transmission of the trait and omitting the indirect trait in order to reduce model complexity. 
We consider a population of $N$ individuals, described by a single trait with a continuous value.
Each generation, $N$ naive individuals, or copiers, choose an individual from the previous generation, or role-models, from which they will copy their trait. Similar to a Wright-Fisher model, we assume non-overlapping generations such that the entire population is replaced in each generation.
The population at time $t$ can be described by $\vec{A}(t)=\big(A_{1}(t), \ldots, A_{N}(t)\big)$ where $A_{i}(t)$ is trait value of individual $i$ at time $t$. We assume the initial population is drawn from a standard normal distribution, $\vec{A}(0) \sim N(0,1)$ .
Cultural transmission is modeled by a function $F$ such that 
\begin{equation}\label{eq:transmission}
A_{i}(t+1) = F_i(\vec{A}(t)) \;.
\end{equation}

\paragraph{Success bias.}
\citet[Ch.~8, p.~247-249]{evolutionBook} describe a blended transmission algorithm by defining $F$ as a weighted average of the traits of all role-models, 
\begin{equation}\label{eq:boydF}
F_i(\vec{A}) = \sum_{j=1}^N G_{i,j}\cdot A_{i,j} \;, 
\end{equation}
where $G_{i,j}$ is the success bias of role-model $j$ in the eyes of copier $i$,
\begin{equation}\label{eq:boydG}
G_{i,j} = \frac{\beta(A_{i,j})}{\sum_{k=1}^{N} \beta(A_{i,k})} \;,
\end{equation}
$A_{i,j}$ is the absolute trait value copier $i$ estimates for role-model $j$ with some error $e_i \sim N(0,\eta^2)$,
\begin{equation}\label{eq:relativeTrait}
A_{i,j} = A_j + e_i,
\end{equation}
and $\beta(\cdot)$ is the bias function that quantifies the success bias of a role-model,
\begin{equation}\label{eq:success_bias}
\beta(A_{i,j}) = b \cdot \exp{\Big(-\frac{(A_{i,j} - \hat{A})^2}{2J}\Big)},
\end{equation} 
with $\hat{A}$ as the arbitrary optimal trait value, and $J$ and $b$ as parameters that control the bias strength.
Therefore, $G_{i,j}$ is a relative success score that copier $i$ assigns to role-model $j$, equivalent to relative fitness in evolutionary-genetic transmission models.

@@

After finding that the DM distribution is a good approximation of the (within-generation) role-model choice process, we turn our attention to the (between-generation) evolutionary dynamics.
We focus on the fixation probability and fixation time of an advantageous phenotype, using a diffusion-equation approximation approach, similar to analyses of population-genetic models \citep{kimura,kimura_average,otto_fixation}.
We are mainly interested in the effect of the bias weight, $\alpha$, which determines the relative effect of success and influence on prestige bias. % TODO ref eq 10?
For simplicity, we do not include role-model estimation error in this analysis. % TODO meaning epsilon=0?
As shown above, transmission in our model is approximately DM distributed (\cref{cor:dirichlet} and \cref{eq:copier_choose}).

We start by finding the expectation and variance of the change in frequency from one generation to the next, which are the drift and diffusion terms of the diffusion equation.\\

\begin{result}[Drift and diffusion terms in a constant environment]
Let $x$ and $x'$ be the frequency of type $\hat{A}$ in a population with $N$ individuals in the current and next generation, and  $\beta$ is the success coefficient of phenotype $A$, $\beta = \beta(A) < \beta(\hat{A}) = 1$.
Then,
\begin{equation}
E[x'-x] \approx x(1-x)(1-\beta) \;, 
\quad
V(x'-x) \approx x(1-x)\Big(\frac{1}{\alpha N + (1-\alpha)}\Big) \;.
\end{equation} 
\end{result}

\begin{proof}
Let $X$ be the number of individuals of type $\hat{A}$ such that $x=X/N$. $X'$ is the number of individuals with $\hat{A}$ in the next generation.
The expected number of individuals is (due to the DM approximation),
\begin{equation}
E[X'] = N  \frac{\alpha_1}{\alpha_1+\alpha_2} \;,
\end{equation}
where $\alpha_1 = \alpha' X$ and $\alpha_2 = \alpha'(N-X)\beta$, from  \cref{eq:binary-model}.
To use frequencies instead of counts, $E[x'] = E[X'/N] = \frac{1}{N}E[X']$.
Putting it together,
\begin{equation}
\begin{split}
E[x'] &= \frac{1}{N}N\frac{\alpha' xN}{\alpha' xN + \alpha' N (1-x)\beta}
	  = \frac{x}{x + (1-x)\beta} \\
	  &= \frac{x}{x + (1-x) -(1-x) + (1-x)\beta}
	  = x \frac{1}{1 -(1-x)(1-\beta)}  \\
	  &= x \big(1 + (1-x)(1-\beta) + o(\beta)\big)
	  = x + x(1-x)(1-\beta) + o(\beta) \;,
\end{split}
\end{equation}
following \citet[p.~253, ch~7.2]{durret} and because $1/(1-y)=1+y+y^2+\ldots$.

By definition, $x$ is constant, so $E[x] = x$.
We therefore have
\begin{equation}\label{eq:expec_freq}
E[x'-x] = E[x'] - E[x] = x(1-x)(1-\beta) + o(1-\beta) \;,
\end{equation}
which gives us the drift term of the diffusion equation.

Using the variance of the DM distribution,
\begin{equation}
V(X') = N\frac{\alpha_1}{\alpha_1+\alpha_2}
\Big(1-\frac{\alpha_1}{\alpha_1+\alpha_2}\Big)
\Big(\frac{N + \alpha_1+\alpha_2}{1+\alpha_1+\alpha_2}\Big) \;.
\end{equation}
Again, we want to use frequencies so we have $V(X'/N) = \frac{1}{N^2}V(x')$.
Putting it together with our model notations,
\begin{equation}
V(x') = \frac{1}{N^2}N\frac{x}{x+(1-x)\beta}\Big(1-\frac{x}{x+(1-x)\beta}\Big)
\Big(\frac{N + \alpha' xN + \alpha' N(1-x)\beta}{1 + \alpha' xN + \alpha' N(1-x)\beta}\Big) \;.
\end{equation}

Following \citet[ch~7.2]{durret}, we assume $\beta \approx 1$, such that
\begin{equation}
\frac{x}{x + (1-x)\beta} \approx x \,
\end{equation}
and for the entire variance expression we get
\begin{equation}
%\begin{split}
V(x') \approx  \frac{1}{N} x(1-x)
\Big(\frac{N + \alpha' xN + \alpha' N - \alpha' xN}{1 + \alpha' xN + \alpha' N - \alpha' xN}\Big)
= x(1-x)\Big(\frac{1 + \alpha'}{1 + \alpha' N}\Big) \;.
%\end{split}
\end{equation}
Now because $x$ is a constant, $V(x) = 0$, and therefore
\begin{equation}\label{eq:var_diff_durret}
V(x'-x) = V(x') - V(x) \approx  x(1-x)(\frac{1 + \alpha'}{1 + \alpha' N}) \;.
\end{equation}
$\alpha'$ is the odds ratio of the bias weight, 
\begin{equation}\label{eq:success_ratio}
\alpha' = \frac{\alpha}{1-\alpha} \;.
\end{equation}
Combining \cref{eq:var_diff_durret} and \cref{eq:success_ratio} we get:
\begin{equation}\label{eq:const_var}
\begin{split}
V(x'-x) \approx x(1-x)\Big(\frac{1 + \frac{\alpha}{1-\alpha}}{1 + \frac{\alpha}{1-\alpha} N}\Big)
% &= x(1-x)(\frac{\frac{1-\alpha+\alpha}{1-\alpha}}{\frac{1-\alpha+\alpha N}{1-\alpha}})\\
% &= x(1-x)(\frac{1}{1- \alpha(1-N)})\\
  = x(1-x)(\frac{1}{\alpha N + (1-\alpha)}) \;.
\end{split}
\end{equation}
This gives the diffusion term of the diffusion equation, which completes the proof.
\end{proof}

This analysis gives a surprising result relating the parameters $\alpha$ and $\beta$ to parameters of the classical Wright-Fisher model from population genetics:
the selection coefficient $s$, a measure of the effect of natural selection on the change in frequency of genotypes, and the effective population size, $N_e$, a measure of the effect of random genetic drift on the change in frequency of genotypes. 
In a diffusion-equation approximation of the classical Wright-Fisher model, the expectation and variance of the change in frequency are $E[x'-x]=x+x(1-x)s+o(s)$ and $V[x'-x]=x(1-x)/N_e$ \citep[eq.~7]{kimura}. 
Therefore, we have the following result.\\

\begin{result}[Effective selection coefficient and population size]\label{res:selection_coef}
The effective selection coefficient $s$ and effective population size $N_e$ can be written in terms of the success coefficient $\beta$ (\cref{eq:success_bias}), the bias weight $\alpha$ (\cref{eq:prestige}), and the population size $N$ as 
\begin{equation}
s=1-\beta(A), \quad N_e=\alpha N + (1-\alpha) \;.
\end{equation}
\end{result}

Using our effective selection coefficient, $1-\beta$, and effective population size, $N_e$, with the population-genetics fixation probability approximation given by \citet[eq.~8]{kimura}, we get the following result.\\

\begin{result}[Fixation probability]
The fixation probability is approximately
\begin{equation}\label{eq:kimura_p}
\pi = \frac{1-e^{-2(1-\beta)N_e x}}{1-e^{-2(1-\beta)N_e}}
\end{equation}
where $x$ is the initial frequency of the advantageous phenotype $\hat{A}$.
\end{result}

Similarly, we can use $1-\beta$ and $N_e$ in the population-genetics fixation time approximation given by \citet[eq.~8]{kimura_average}.\\

\begin{result}[Fixation time approximation]
\begin{equation}\label{eq:kimura_t}
T = \frac{1-\pi}{1-\beta}\int_0^x\frac{e^{2(1-\beta) \xi}-1}{\xi(1-\xi)}d\xi+ \frac{\pi}{1-\beta}\int_x^1\frac{1-e^{-2(1-\beta)(1-\xi)}}{\xi(1-\xi)}d\xi
\end{equation}
% TODO double check this, because eq 17 in Kimura 1969 looks different.
% in particular, I don't see Ne here...
\end{result}
Note that these integrals cannot be solved in closed form, so we can only estimate them numerically.

\paragraph{Numerical validation.}
To validate our analysis we performed multiple simulations comparing our binary model with the classic Wright-Fisher model, using different $\alpha$ and $\beta$ each time, and using the corresponding values of $s=1-\beta$ and $N_e=\alpha N + (1-\alpha)$ for the Wright-Fisher simulations.
%In \cref{fig:var_alpha} we changed $\alpha$ (and $N_e$ accordingly) and used a constant $\beta$.
%In \cref{fig:var_beta} we changed $\beta$ and used a constant $\alpha$.
We find see that the two models have similar dynamics, and both are well approximated the the approximations in \cref{eq:kimura_p,eq:kimura_t}.

\begin{figure}[h]
% TODO 
% capital letters in panel labels
% bigger font
    \includegraphics[width=\linewidth]{../figures/binary/kimura_var_alpha.pdf}
  \caption{\textbf{Fixation probability and time approximation for varying bias weight.}
  Fixation probability (\textbf{a}) and time (\textbf{b}) for varying bias weight on the bottom x-axis. The corresponds effective population size is on the top x-axis.
  The approximation (black) fits DM simulations results (green) and Wright-Fisher simulation results (orange) when using \cref{eq:kimura_p}.
  Fixation probability (a) is limited at approximately $2s$ (blue), as in the classic WF model.
  Markers are averages of $5,000$ simulations, error bars show 95\% confidence intervals.
   Here, Population size, $N=1,000$; phenotype values, $\hat{A}=1$, $A=0.7$; success coefficient, $1-\beta=s=0.044$.}
  \label{fig:var_alpha}
\end{figure}

\begin{figure}[h]
    \includegraphics[width=\linewidth]{../figures/binary/kimura_var_beta.pdf}
  \caption{\textbf{Fixation probability and time approximation for varying success coefficient.}
  % TODO
  % merge with previous figure; see comments there.
   The approximation (black) fits DM simulations results (green) and Wright Fisher simulation results (orange) when using \cref{eq:kimura_p}.
  Markers for average values across $5,000$ simulations, error bars for 75\% confidence intervals.
 Here, Population size, $N=1,000$; phenotype values, $\hat{A}=1$, $A=a\cdot \hat{A}, a \in [0.01,0.99]$; bias weight, $\alpha=0.01$.}
  \label{fig:var_beta}
\end{figure}

After finding a good approximation for our model in a constant environment, when the favorable trait is always $\hat{A}$, we continued to find an approximation for our model in a changing environment. 
Following \citet{changeEnv}, we find the following expressions for the expectation and variance of the change in frequency between $t$ generations.\\

\begin{result} [Drift and diffusion terms in a changing environment]\label{res:ch_expected}
Let $x$ be the initial frequency of the invading phenotype and $X_t$ is the number of individuals with the phenotype at time $t$.
Then,
\begin{equation}\label{eq:ch_expected_and_var}
E[X_t/N - x] \simeq x(1-x) S_t / N \;, 
\quad
V(X_t/N ) \simeq x(1-x) t/N_e \;, % TODO shouldn't it be V[X_t/N - x] ?? or something like that? with square brackets and with "- something" ?
\end{equation}
where $S_t=\sum\limits_{i=1}^{t} N (1-\beta_t)$.
\end{result}

\begin{proof}
Let $s_t=N(1-\beta_t)$, and $S_n=\sum\limits_{i=1}^n s_i$, where $\beta_t$ is $\beta(A)$ at generation $t$.
We prove by induction both terms in \cref{eq:ch_expected_and_var}.
From \cref{eq:expec_freq} we know that
\begin{equation}\label{eq:ch_1}
%\begin{split}
E\left[\frac{X_{t+1}}{N}-\frac{X_t}{N}\bigg|X_t\right] 
= \frac{X_t}{N}\left(1-\frac{X_t}{N}\right)(1-\beta_{t+1}) 
= \frac{1}{N}\frac{X_t}{N}\left(1-\frac{X_t}{N}\right)s_{t+1} \;.
%\end{split}
\end{equation}
Also note that using the definition of $V(y)=E[y^2]-(E[y])^2$
\begin{equation}
\begin{split}
E\left[\frac{X_t}{N}\left(1-\frac{X_t}{N}\right)\right] 
&= E\left[\frac{X_t}{N}-\left(\frac{X_t}{N}\right)^2\right] 
= E\left[\frac{X_t}{N}\right] - E\left[\left(\frac{X_t}{N}\right)^2\right] \\
&= E\left[\frac{X_t}{N}\right] - V\left(\frac{X_t}{N}\right) - \left(E\left[\frac{X_t}{N}\right]\right)^2 \;.
\end{split}
\end{equation}

We can now use the induction assumption of $V(\frac{X_t}{N})$ to get
\begin{equation}
\begin{split}
E\left[\frac{X_t}{N}\left(1-\frac{X_t}{N}\right)\right] 
&\simeq E\left[\frac{X_t}{N}\right]\left(1-E\left[\frac{X_t}{N}\right]\right)-\frac{1}{N_e}tx(1-x) \;.
\end{split}
\end{equation}
From \cref{eq:ch_1} we know that
\begin{equation}
\begin{split}
E\left[\frac{X_{t+1}}{N}-\frac{X_t}{N}\right] 
&= \frac{1}{N}s_{t+1}E\left[\frac{X_t}{N}\left(1-\frac{X_t}{N}\right)\right] 
\simeq \frac{1}{N}s_{t+1}\left(E\left[\frac{X_t}{N}\right]\left(1-E\left[\frac{X_t}{N}\right]\right) - \frac{1}{N_e}tx(1-x)\right) \\
&\simeq \frac{1}{N}s_{t+1}\cdot E\left[\frac{X_t}{N}\right]\left(1-E\left[\frac{X_t}{N}\right]\right) - \frac{1}{N_e N}s_{t+1}tx(1-x) \;.
\end{split}
\end{equation}
Now we omit $O(\frac{1}{Ne\cdot N})$ and get
\begin{equation}\label{eq:ch_2}
E\left[\frac{X_{t+1}}{N}-\frac{X_t}{N}\right] \simeq \frac{1}{N}s_{t+1}\cdot E\left[\frac{X_t}{N}\right]\left(1-E\left[\frac{X_t}{N}\right]\right) \;.
\end{equation}
